# Capital Cities (Urban Mobility)

## Descriptive Statistics 

```{r initialization, eval = TRUE, echo = FALSE, results = hide}
## initialize and load required packages
library(RPostgreSQL)
library(ggplot2)
library(scales)

## initialize database connection and psql helper functions
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, 
                 host = "localhost", 
                 port = "1111", 
                 user = "postgres", 
                 password = "postgres", 
                 dbname = "pflow")

source("psqlHelper.R")

## initialize ggplot
theme_set(theme_bw())
```

### Sample Populations

This section covers the descriptive statistics of the four sample populations.

```{r loadVariableLabels, eval = TRUE, echo = FALSE, results = hide}
# load variable labels
labelsSex <- read.table("labels_all_sex.txt", sep = "\t")
labelsAge <- read.table("labels_all_age.txt", sep = "\t")
labelsOccupC <- read.table("labels_all_occupc.txt", sep = "\t")
labelsDhk09Occup <- read.table("labels_dhk09_occup.txt", sep = "\t")
labelsHni04Occup <- read.table("labels_hni04_occup.txt", sep = "\t")
labelsJkt02Occup <- read.table("labels_jkt02_occup.txt", sep = "\t")
labelsMnl96Occup <- read.table("labels_mnl96_occup.txt", sep = "\t")
```

#### Dhaka 2009

```{r loadDhk09Person, eval = TRUE, echo = FALSE}
## dhk09
dhk09.person <- psqlGetTable(con, "dhk09", "person")
# remove unnecessary columns (padd, magfac, magfac2)
dhk09.person <- dhk09.person[, c("pid", "sex", "age", "occup", "occupc")]
# factorize pid
dhk09.person$pid <- factor(dhk09.person$pid)
# remove NA sex (=0) and factorize
dhk09.person <- dhk09.person[dhk09.person$sex != 0, ]
dhk09.person$sex <- factor(dhk09.person$sex, 
                           levels = c(1, 2), 
                           labels = labelsSex[, 2])
# factorize age (ordered)
dhk09.person$age <- factor(dhk09.person$age, 
                           levels = c(0:17), 
                           ordered = TRUE, 
                           labels = labelsAge[, 2])
# factorize occupation
dhk09.person$occup <- factor(dhk09.person$occup, 
                             levels = c(0:8), 
                             labels = labelsDhk09Occup[, 2])
# factorize occupation class
dhk09.person$occupc <- factor(dhk09.person$occupc, 
                              levels = c(1:5, 99), 
                              labels = labelsOccupC[, 2])
```
The data were captured on October 1-2, 2009 with a sample size of `r format(nrow(dhk09.person), big.mark=",")` people.

Data set description:
```{r descrDhk09Person}
names(dhk09.person)
str(dhk09.person)
levels(dhk09.person$sex)
levels(dhk09.person$age)
levels(dhk09.person$occup)
levels(dhk09.person$occupc)
```

```{r plotHistDhk09Person, eval = TRUE, echo = FALSE}
ggplot(dhk09.person, aes(sex)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Dhaka 2009, Sex", x = "Sex", y = "Count")
ggplot(dhk09.person, aes(age)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Dhaka 2009, Age", x = "Age", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplot(dhk09.person, aes(occup)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Dhaka 2009, Occupation", x = "Occupation", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplot(dhk09.person, aes(occupc)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Dhaka 2009, Occupation Group", x = "Occupation", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


#### Hanoi 2004

```{r loadHni04Person, eval = TRUE, echo = FALSE}
## hni04
hni04.person <- psqlGetTable(con, "hni04", "person")
# remove unnecessary columns (padd, magfac, magfac2)
hni04.person <- hni04.person[, c("pid", "sex", "age", "occup", "occupc")]
# factorize pid
hni04.person$pid <- factor(hni04.person$pid)
# factorize sex
hni04.person$sex <- factor(hni04.person$sex, 
                           levels = c(1, 2), 
                           labels = labelsSex[, 2])
# factorize age (ordered)
hni04.person$age <- factor(hni04.person$age, 
                           levels = c(0:17), 
                           ordered = TRUE, 
                           labels = labelsAge[, 2])
# factorize occupation
hni04.person$occup <- factor(hni04.person$occup, 
                             levels = c(0:16, 99), 
                             labels = labelsHni04Occup[, 2])
# factorize occupation class
hni04.person$occupc <- factor(hni04.person$occupc, 
                              levels = c(1:5, 99), 
                              labels = labelsOccupC[, 2])
```
The data were captured on October 1-2, 2004 with a sample size of `r format(nrow(hni04.person), big.mark=",")` people.

Data set description:
```{r descrHni04Person}
names(hni04.person)
str(hni04.person)
levels(hni04.person$sex)
levels(hni04.person$age)
levels(hni04.person$occup)
levels(hni04.person$occupc)
```

```{r plotHistHni04Person, eval = TRUE, echo = FALSE}
ggplot(hni04.person, aes(sex)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Hanoi 2004, Sex", x = "Sex", y = "Count")
ggplot(hni04.person, aes(age)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Hanoi 2004, Age", x = "Age", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplot(hni04.person, aes(occup)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Hanoi 2004, Occupation", x = "Occupation", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplot(hni04.person, aes(occupc)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Hanoi 2004, Occupation Group", x = "Occupation", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


#### Jakarta 2002

```{r loadJkt02Person, eval = TRUE, echo = FALSE}
## jkt02
jkt02.person <- psqlGetTable(con, "jkt02", "person")
# remove unnecessary columns (padd, magfac, magfac2)
jkt02.person <- jkt02.person[, c("pid", "sex", "age", "occup", "occupc")]
# factorize pid
jkt02.person$pid <- factor(jkt02.person$pid)
# factorize sex
jkt02.person$sex <- factor(jkt02.person$sex, 
                           levels = c(1, 2), 
                           labels = labelsSex[, 2])
# factorize age (ordered, no 0)
jkt02.person$age <- factor(jkt02.person$age, 
                           levels = c(0:17), 
                           ordered = TRUE, 
                           labels = labelsAge[, 2])
# factorize occupation
jkt02.person$occup <- factor(jkt02.person$occup, 
                             levels = c(0:17, 99), 
                             labels = labelsJkt02Occup[, 2])
# factorize occupation class
jkt02.person$occupc <- factor(jkt02.person$occupc, 
                              levels = c(1:5, 99), 
                              labels = labelsOccupC[, 2])
```
The data were captured on October 1-2, 2002 with a sample size of `r format(nrow(jkt02.person), big.mark=",")` people.

Data set description:
```{r descrJkt02Person}
names(jkt02.person)
str(jkt02.person)
levels(jkt02.person$sex)
levels(jkt02.person$age)
levels(jkt02.person$occup)
levels(jkt02.person$occupc)
```

```{r plotHistJkt02Person, eval = TRUE, echo = FALSE}
ggplot(jkt02.person, aes(sex)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Jakarta 2002, Sex", x = "Sex", y = "Count")
ggplot(jkt02.person, aes(age)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Jakarta 2002, Age", x = "Age", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplot(jkt02.person, aes(occup)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Jakarta 2002, Occupation", x = "Occupation", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplot(jkt02.person, aes(occupc)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Jakarta 2002, Occupation Group", x = "Occupation", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


#### Metro Manila 1996

```{r loadMnl96Person, eval = TRUE, echo = FALSE}
## mnl96
mnl96.person <- psqlGetTable(con, "mnl96", "person")
# remove unnecessary columns (padd, magfac, magfac2)
mnl96.person <- mnl96.person[, c("pid", "sex", "age", "occup", "occupc")]
# factorize pid
mnl96.person$pid <- factor(mnl96.person$pid)
# factorize sex
mnl96.person$sex <- factor(mnl96.person$sex, 
                           levels = c(1, 2), 
                           labels = labelsSex[, 2])
# factorize age (ordered, no 0)
mnl96.person$age <- factor(mnl96.person$age, 
                           levels = c(0:17), 
                           ordered = TRUE, 
                           labels = labelsAge[, 2])
# factorize occupation
mnl96.person$occup <- factor(mnl96.person$occup, 
                             levels = c(1:14, 99), 
                             labels = labelsMnl96Occup[, 2])
# factorize occupation class
mnl96.person$occupc <- factor(mnl96.person$occupc, 
                              levels = c(1:5, 99), 
                              labels = labelsOccupC[, 2])
```
The data were captured on October 1-2, 1996 with a sample size of `r format(nrow(mnl96.person), big.mark=",")` people.

Data set description:
```{r descrMnl96Person}
names(mnl96.person)
str(mnl96.person)
levels(mnl96.person$sex)
levels(mnl96.person$age)
levels(mnl96.person$occup)
levels(mnl96.person$occupc)
```

```{r plotHistMnl96Person, eval = TRUE, echo = FALSE}
ggplot(mnl96.person, aes(sex)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Metro Manila 1996, Sex", x = "Sex", y = "Count")
ggplot(mnl96.person, aes(age)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Metro Manila 1996, Age", x = "Age", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplot(mnl96.person, aes(occup)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Metro Manila 1996, Occupation", x = "Occupation", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplot(mnl96.person, aes(occupc)) +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Dhaka 2009, Occupation Group", x = "Occupation", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


# Final Steps

It's always a good idea to clean up after you're done...

```{r cleanup, eval = TRUE, echo = FALSE, results = hide}
## clean up database connection and database driver
dbDisconnect(con)
dbUnloadDriver(drv)
```
